Improving phase-matching of FWM with temperature

### 2015-09-19 latex/article.tex
computational model section & other

\documentclass[12pt,a4paper,twoside]{article}
%\documentclass[12pt,a4paper,twocolumn,twoside]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{parcolumns}
\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage[final]{pdfpages} %serve per aggiungere altre pagine pdf al file
\usepackage{siunitx}

@@

\pagenumbering{roman}	% numeration in roman numbers
\begin{document}

@@

The study is focused on a silicon-based waveguide with a silicon ($\mathrm{Si}$) core of rectangular section and a silica ($\mathrm{Si0}_2$) cladding.
The purpose of the waveguide is to generate light through a partial degenerate Four-Wave Mixing phenomenon.

@@

The general solution of Eq. (\ref{eq_helm_P}) is a transverse electromagnetic (TEM) wave, whose electric and magnetic fields are perpendicular to the propagation direction.
Without losing generality, we could assume $\hat{z}$ as the propagation direction and electric field of the form:
\begin{equation}
	\textbf{E} = \frac{1}{2}\hat{x}\,E\,\mathrm{exp}[\mathrm{i}(kz-\omega t)] + c.c.
	\label{eq_wave}
\end{equation}
with $E$ amplitude of the electric field, $k = 2\pi/\lambda$ the wavevector and $\omega = 2\pi f$ the angular frequency. $\lambda$ and $f$ are relatively the wavelength and the frequency of the electromagnetic wave.
The magnetic field has a similar equation.

@@

In order to be efficient, the parametric process needs to verify both the conservation of energy and momentum.
$$\omega_1 + \omega_2 = \omega_3 + \omega_4$$
$$\Delta k = \beta_3 + \beta_4 - \beta_1 - \beta_2$$
****
$$\Omega_S = \omega_1 + \omega_3 = \omega_4 - \omega_1$$
where we assume $\omega_3 < \omega_4$.
$$\Omega_S = |\omega_P - \omega_I| = |\omega_S - \omega_P|$$
****

parlare di $L_{coh}$

@@

\section{Computational models}
\subsection{Simulation}
To understand the behaviour of the waveguide, it is necessary to first know how waves of different frequency are transmitted through the waveguide.
We chose to simulate the modes with temperature and frequency as a parameters.

Our aim was to develop a model to work with an infrared laser pump with $\lambda_{p} = \SI{1.55}{\um}$.
We narrowed our simulations to a frequency gap centered in $\omega_p = 2\pi \nu_p = 2\pi c / \lambda_{p}$.
The gap chosen was precisely $[\SI{1.45}{\um} ,\, \SI{1.65}{\um}]$ or the equivalent $[\SI{182}{\THz} ,\, \SI{207}{\THz}]$.

The reason for this choice is that silicon is transparent at these wavelength and also because of the availability of lasers that emit at the specific wavelength.

\vspace{18pt}
****
\vspace{12pt}

Our data output was therefore the field distribution and effective index for a number of modes for each parametric configuration.

The next step was to classify these data depending on the order and type (TE or TM) of the modes.
We achieved this by analyzing the maxima of the field distributions.
We obtained then a well organized data matrix filled with the value of effective index for each combination of parameters.

\subsection{Data processing}
Starting from raw data consisting of the effective index for each parametric configuration, we fitted them with a polinomial function of frequency $\omega = 2\pi \nu$ and temperature $T$, obtaining:
$$n_{eff} = f_{mt,\,mo} \left( \omega, T \right)$$
where $mt$ and $mo$ means respectively \textit{mode type} (TE or TM) and \textit{mode order} (1, 2, 3, ...)

\vspace{18pt}
************************************

\subsubsection*{Order combination}

@@

\subsubsection*{Conservation of energy and momentum}
Starting from the data of the effective index of each order at different wavelengths and temperatures, we evaluated the phase mismatch of all sets.
The evaluation was made on a grid of temperature and signal wavelength.

@@

\subsubsection*{Selection of sets}
Once evaluated all the points on the grid for all the sets available, we selected only the interesting combinations.
The critical parameter on which the choice was based on is the coherence length: only those sets which had a maximum value of coherence length greater than a supposed sample length $L_{sam} = $\SI{1}{\cm} was chosen.
$$L_{coh} = \frac{2\pi}{|\Delta k|} \geq L_{sam}$$
Again, we had a list of the only combination interesting for our purpose.

@@

\subsubsection*{Data aggregation}
After the selection of sets, we evaluated the dependence of each maximum in temperature.
\section{Prototyping}

\subsection{Heater configurations}
In order to achieve a greater effect from the different temperature, we tried to move the heater from over the center of the core to the side.
This operation was intended to increase the difference of behaviour between low and high orders enabling thus the creation of new peaks in the coherence length.
\subsection{Different geometries}

\section{Conclusion}

\cleardoublepage
\begin{thebibliography}{9}
\bibitem{agrawal} G. Agrawal. Nonlinear Fiber Optics, 5th edition. 2013
\bibitem{saleh} B. E. A. Saleh, M. C. Teich. Fundamentals of photonics, 2nd edition. 2007
\end{thebibliography}
\end{document}
% Ref. \cite{agrawal}

>>>>

\documentclass[12pt,a4paper,twoside]{article}
%\documentclass[12pt,a4paper,twocolumn,twoside]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{parcolumns}
\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage[final]{pdfpages} %serve per aggiungere altre pagine pdf al file
\usepackage{siunitx}
\usepackage{booktabs}	% better tables
\usepackage{multirow}

@@

\makeatletter
\renewcommand\paragraph{
   \@startsection{paragraph}{4}{\z@}%
   {-3.25ex\@plus -1ex \@minus -.2ex}%
   {1.5ex \@plus .2ex}%
   {\normalfont\normalsize\bfseries}}
\makeatother

\pagenumbering{roman}	% numeration in roman numbers

\begin{document}

@@

The study is focused on a silicon-based waveguide with a silicon ($\mathrm{Si}$) core of rectangular section and a silica ($\mathrm{SiO}_2$) cladding.
The purpose of the waveguide is to generate light through a partial degenerate Four-Wave Mixing phenomenon.

@@

The general solution of Eq. (\ref{eq_helm_P}) is a transverse electromagnetic (TEM) wave, which electric and magnetic fields are perpendicular to the propagation direction.
Without losing generality, we could assume $\hat{z}$ as the propagation direction and electric field of the form:
\begin{equation}
	\textbf{E} = \frac{1}{2}\hat{x}\,E\,\mathrm{exp}[\mathrm{i}(kz-\omega t)] + c.c.
	\label{eq_wave}
\end{equation}
with $E$ amplitude of the electric field, $k = 2\pi/\lambda$ the wavevector and $\omega = 2\pi f$ the angular frequency. $\lambda$ and $f$ are relatively the wavelength and the frequency of the electromagnetic wave.
The magnetic field has a similar equation.

@@

In order to be efficient, the parametric process needs to verify both the conservation of energy and momentum.
$$\omega_1 + \omega_2 = \omega_3 + \omega_4$$
$$\Delta k = \beta_3 + \beta_4 - \beta_1 - \beta_2$$
****
$$\Omega_S = \omega_1 + \omega_3 = \omega_4 - \omega_1$$
where we assume $\omega_3 < \omega_4$.
$$\Omega_S = |\omega_P - \omega_I| = |\omega_S - \omega_P|$$

@@

\section{Computational models}
To understand the behaviour of the waveguide, it is necessary to first know how waves of different frequency are transmitted through the waveguide.
We chose to simulate the modes with temperature and frequency as a parameters.

Our aim was to develop a model to work with an infrared laser pump with $\lambda_{p} = \SI{1.55}{\um}$.
The reason for this choice is that silicon is transparent at these wavelength and also because of the availability of lasers that emit at the specific wavelength.

We narrowed our simulations to a frequency gap centered in $\omega_p = 2\pi \nu_p = 2\pi c / \lambda_{p}$.
The gap chosen was precisely $[\SI{1.45}{\um} ,\, \SI{1.65}{\um}]$ or the equivalent $[\SI{182}{\THz} ,\, \SI{207}{\THz}]$.
The temperature gap, on the other hand, was chosen to be $[\SI{0}{\K} - \SI{400}{\K}]$ over ambient temperature $T_{amb} = \SI{293.15}{\K} = \SI{20}{\celsius}$, that is $[\SI{293.15}{\K} - \SI{693.15}{\K}]$.

\subsection{Simulation}
To obtain definite solutions, we had to assign the materials and a geometry to the problem first.

\begin{figure}[ht]
	\centering
	\includegraphics[width=.55\textwidth]{geometry.png}
	\label{fig_wg_Section}
	\caption{Section of the rectangular waveguide}
\end{figure}

We modelized a dielectric waveguide with a silicon (Si) core \SI{3}{\um} wide and \SI{.25}{\um} high.
The cladding was set to be silica (SiO2) and coated the top of the core with a thickness of \SI{.35}{\um}.
For both materials was defined a refractive index dependent on the wavelength $n=n(\lambda)$, as given by the empirical relationship called \textit{Sellmeier equation}.
\begin{equation}
	n^2(\lambda) = 1+ \sum_i \chi_{0i}\frac{\lambda^2}{\lambda^2-\lambda^2_i}
	\label{eq_sellmeier}
\end{equation}
where the weights are defined experimentally and vary depending on the material.

We positioned a thermal reservoir, a ``\textit{heater}", \SI{0.3}{\um} wide at the top of the cladding, exactly over the center of the waveguide core and set to mantain a specific temperature in the gap, given by the configuration chosen.
To 
define the density $\rho$ in [\si{\kg\m^3}], the thermal conductivity $\kappa_{TH}$ in [\si{\W\per\m\per\K}] and the heat capacity $C_P$ in [\si{\J\per\kg\per\K}] of the different materials.

\begin{table}[ht]
\centering
  \begin{tabular}{lcccc}  
    \toprule
		& \multicolumn{3}{c}{Properties} \\
    \cmidrule(r){2-5}
    	\multirow{2}{*}{Materials}    & $\rho$ & $\kappa_{TH}$ & $C_P$ & TOC\\
  		& [\si{\kg\m^3}] & [\si{\W\per\m\per\K}] & [\si{\J\per\kg\per\K}] & [$10^{-4}$\si{\per\K}] \\
    \midrule
    Silicon (Si)			& 2329	& 130	& 700	& 1.87\\
    Silica (SiO$_2$)		& 2203	& 1.38	& 703	& 0\\
    \bottomrule
  \end{tabular}
  \label{tab_materials}
  \caption{Thermal properties of silicon and silica glass}
\end{table}

Moreover, in order to be influenced by the different temperatures, we had to add to the definition of refractive index of the silicon a term dependent on the temperature.
For our purpose this dependence can easily be considered linear:
\begin{equation}
n = n(\lambda)+\mathrm{TOC}\cdot T
\end{equation}
where $\mathrm{TOC}|_{\mathrm{Si}} = \SI{1.87e-4}{\per\K}$\cite{n_db} is called Termo-Optic Coefficient. The coefficient of silica glass is negligible and coefficient was therefore considered zero $\mathrm{TOC}|_{\mathrm{SiO}_2} \approx 0$.

\subsubsection{Numerical solution}
Because the degree of freedom of the problem were too many, we resorted to automated algorithms which gave us a numerical, and therefore approximate, solution.

\paragraph{Heat transfer: stationary solution}
The first step was solving the equation for the heat transfer to obtain a stationary solution for each temperature in the gap.
That means to know the local temperature for each position in the core and in the cladding, for each initial setup.

\paragraph{EM mode analysis}

The second step was to find the different modes of the waveguide for each set of parameters, such as temperature and wavelength.

We emploied an automated algorithm to solve, in the frequency domain, the Helmholtz equation on a fine grid in the waveguide section.
Each point of evaluation took into account the effective index of the material, slightly changed by the local temperature obtained by the stationary solution.

This algorithm was set to find the eight modes which effective index $n_{eff}$ was closest to $n_{eff}=4$ in absolute value.
\subsubsection{Data output}
We obtained the field distribution and a series of parameters as our data output for each parametric configuration.
In particular the parameters were the effective index $n_{eff}$, the effective area $A_{eff}$ and the nonlinear parameter $\gamma$, for each mode and for each wavelength and temperature.

\subsection{Data classification}
The next step was to classify these data, depending on the type (TE or TM) and order (1,2,3...) of the modes.
We achieved this by analyzing a sample taken at $y=0$ of the field distributions.
The ratio between the maxima values of the x-component and y-component of the electric field allowed to distinguish the mode type: TE or TM.
Those fields which had the maximum component in the y direction were categorized as TE modes while those which had the maximum component in the x direction as TM modes.
The number of maxima of the modulus of the field showed us the order of the mode.

We obtained then a well organized data matrix filled with the value of effective index $n_{eff}$, effective area $A_{eff}$ and nonlinear parameter $\gamma$ for each combination of parameters.

\subsection{Data processing}

\subsubsection{Polynomial fit}
Starting from raw data consisting of the effective index for each parametric configuration, we fitted them with a polinomial function of frequency $\omega = 2\pi \nu$ or $\lambda$ and temperature $T$, obtaining:
$$n_{eff} = f_{mt,\,mo} \left( \omega, T \right) = f'_{mt,\,mo} \left( \lambda, T \right)$$
where $mt$ and $mo$ means respectively \textit{mode type} (TE or TM) and \textit{mode order} (1, 2, 3, ...)

Specifically, we used a polynomial expression of fifth order in wavelength and of second order in temperature ``\texttt{poly52}".

\subsubsection{Order combination}

@@

\subsubsection{Conservation of energy and momentum}
Starting from the data of the effective index of each order at different wavelengths and temperatures, we evaluated the phase mismatch of all sets.
The evaluation was made on a grid of temperature and signal wavelength.

@@

\subsubsection{Selection of sets}
Once evaluated all the points on the grid for all the sets available, we selected only the interesting combinations.
The critical parameter on which the choice was based on is the coherence length: only those sets which had a maximum value of coherence length greater than a supposed sample length $L_{sam} = $\SI{1}{\cm} was chosen.
$$L_{coh} = \frac{2\pi}{|\Delta k|} \geq L_{sam}$$
Again, we had a list of the only combination interesting for our purpose.

@@

\subsubsection{Data aggregation}
After the selection of sets, we evaluated the dependence of each maximum on temperature.

\subsection{Second heater configuration}
In order to achieve a greater effect from the different temperature, we tried to move the heater from over the center of the core to the side.
This operation was intended to increase the difference of behaviour between low and high orders enabling thus the creation of new peaks in the coherence length.
\subsection{Different geometries}

\section{Conclusions}
\subsection{Improvements}
\subsection{Results}

\cleardoublepage
\begin{thebibliography}{9}
\bibitem{agrawal} G. Agrawal. Nonlinear Fiber Optics, 5th edition. 2013
\bibitem{saleh} B. E. A. Saleh, M. C. Teich. Fundamentals of photonics, 2nd edition. 2007
\bibitem{n_db} http://refractiveindex.info/
\end{thebibliography}
\end{document}
% Ref. \cite{agrawal}
